---
title: "Factors of factorials"
author: "Clare Gibson"
date: "2025-07-06"
format:
  gfm:
    toc: true
    toc-depth: 2
---

# Introduction
In maths, there is a concept called **factorial** which works like this:

$$
\begin{align}
2!&=2\times1=2\\
3!&=3\times2\times1=6\\
4!&=4\times3\times2\times1=24
\end{align}
$$

The factors of an integer are the integers that divide equally into it. The factors of $2!$ are 1 and 2. There are two factors.

$$
\begin{align}
\textsf{Factors of 2! (=2):}\\
\quad 1, 2\\
\\
\textsf{Factors of 3! (=6):}\\
\quad 1,2,3,6\\
\\
\textsf{Factors of 4! (=24):}\\
\quad 1,2,3,4,6,8,12,24
\end{align}
$$

The relationship between factorials and factors looks like this:

|$n$|$n!$|number of factors of $n$|
|:---|:---|:---|
|1   |1   |1   |
|2   |2   |2   |
|3   |6   |4   |
|4   |24  |8   |

# Questions

- How many factors has $5!$ got?
- How many factors has $6!$ got?
- What pattern do you notice?

# Analysis

In the `src/factors-of-factorials.R` script I have written a function called `get_factors()` which takes an integer as input and returns a list of the factors of that integer. I will load the script so that I can use this function.

```{r}
#| label: setup
#| warnings: false
#| output: false

library(here)
library(tidyverse)
library(knitr)

source(here("src/factors-of-factorials.R"))
```

I can test the function using one of the integers we already outlined above.

```{r}
#| label: test-function

get_factors(6)
```

Next, I will build a dataframe to view the relationships for a greater set of numbers (from 0 to 10):

```{r}
#| label: build-dataframe

factors_of_factorials <- tibble(n = 1:10) |> 
  mutate(
    n_factorial = factorial(n),
    factors_of_n_factorial = map(n_factorial, get_factors),
    total_factors = map_int(factors_of_n_factorial, length)
  )

kable(factors_of_factorials)
```

# Plots

Finally I can plot the results on a line chart.

```{r}
#| label: line-chart

factors_of_factorials |> 
  ggplot(aes(x = n, y = total_factors)) +
  geom_line()
```

# Conclusion

The line chart shows that the relationship between an integer $n$ and the number of factors that its factorial has is exponentially increasing.